---
title: "Mapping Economic Impact: A Global View of Child Deprivation and Prosperity"
author: "Manthan Tiwari"
format: html
editor: visual
execute: 
  echo: false
  warning: false
  message: false
---

# Abstract:

This report explores the relationship between economic indicators and child deprivation across various countries. Utilizing data from two key sources—unicef_indicator_1.xlsx and unicef_metadata.xlsx—the analysis is presented through a series of visualizations that illustrate global patterns and trends. These insights aim to inform policymakers and stakeholders, providing a data-driven foundation for targeted interventions to improve child welfare globally.

# Introduction:

Child deprivation is a critical issue that affects future generations' educational outcomes, health, and overall well-being. Understanding the economic factors that correlate with child deprivation can help tailor effective policies. This report uses comprehensive datasets from UNICEF to examine the linkage between GDP per capita and child deprivation rates worldwide.

# Data Description:

The analysis is based on two primary data sources:

unicef_indicator_1.xlsx: Contains data on child deprivation rates, detailing the percentage of children suffering from educational, nutritional, or welfare inadequacies in various countries. unicef_metadata.xlsx: Provides broader economic and demographic indicators such as GDP per capita, population totals, and life expectancy.

## Visualization 1: World Map Chart

#### Narrative:

This world map vividly illustrates the prevalence of **child deprivation across different nations**, highlighting regions with the **highest and lowest percentages**. By visualizing **deprivation data geographically,** we can immediately identify **global patterns and hotspots** where interventions are most needed. This map sets the stage for a deeper exploration into what drives these patterns, particularly the role of economic factors.

```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("sf")
install.packages("plotly")
```

```{r}
source("main.r")
```

```{r}
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")

# Merge world map data with our data
world_map <- left_join(world, combined_data, by = c("iso_a2" = "alpha_2_code"))

# Plot
ggplot(data = world_map) +
  geom_sf(aes(fill = Child_Deprivation)) +
  labs(title = "Global Child Deprivation Rates", fill = "Deprivation %") +
  theme_minimal()
```

## Visualization 2: Bar Chart

#### Title: Top 10 Countries by GDP per Capita in 2017

#### Narrative:

Focusing on economic prosperity, this bar chart ranks the **top 10 countries by GDP per capita in 2017.** It reveals a snapshot of **economic health in the wealthiest nations**, providing a stark contrast to the child deprivation rates seen in the map. This comparison raises questions about the distribution of wealth and its impact on societal well-being, prompting us to look closer at individual countries where economic growth may not equate to reduced child deprivation.

```{r}
selected_countries <- combined_data %>%
  filter(year == 2017) %>%
  arrange(desc(GDP_per_capita)) %>%
  top_n(10, GDP_per_capita)

ggplot(selected_countries, aes(x = reorder(Country, GDP_per_capita), y = GDP_per_capita)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Countries by GDP per Capita in 2017", x = "Country", y = "GDP per capita (US$)")
```

## Visualization 3: Scatter-plot with Linear Regression Line

#### Title: Child Deprivation vs. GDP Per Capita (2017)

#### Narrative:

Delving into the relationship between **economic output** and **child welfare**, this scatterplot with a linear regression line examines how **GDP per capita** correlates with child deprivation percentages across countries. The trend line offers a statistical backbone to our observations, highlighting whether higher economic output generally corresponds to lower levels of child deprivation, or if the reality is more complex.

```{r}
plot_data <- combined_data %>%
  filter(year == 2017)

ggplot(plot_data, aes(x = GDP_per_capita, y = Child_Deprivation)) +
  geom_point(aes(color = Country)) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Child Deprivation vs. GDP per Capita (2017)", x = "GDP per Capita", y = "Deprivation %")
```

## Visualization 4: Time-Series Chart

#### Title: GDP Per Capita Over Time in India

#### Narrative:

Zooming into India, this time-series chart tracks GDP per capita over several decades, illustrating the trajectory of economic growth. By examining one country in detail, we can discuss the nuances of economic reforms, development policies, and their timings relative to changes in child welfare metrics. It provides a longitudinal perspective, crucial for understanding long-term impacts.

```{r}
india_data <- combined_data %>%
  filter(Country == "India")

ggplot(india_data, aes(x = year, y = GDP_per_capita)) +
  geom_line() +
  geom_point() +
  expand_limits(y = 0) +
  labs(title = "GDP Per Capita Over Time in India", x = "Year", y = "GDP per capita (US$)")
```

## Visualization 5: Interactive Plots

#### Title: Explore the Data: Interactive Analysis of GDP and Child Deprivation

#### Narrative:

This interactive plot invites users to engage directly with the data, exploring relationships between GDP, child deprivation, and other variables like population size. By interacting with the plot, users can select specific countries or years, dive into the details, and form their own hypotheses about the data. It's a tool for discovery and education, enabling a personal journey through the data where each user can draw unique insights.

```{r}
library(plotly)
plot_ly(data = combined_data, x = ~GDP_per_capita, y = ~Child_Deprivation, type = 'scatter', mode = 'markers',
        marker = list(size = 10, color = 'rgba(255, 182, 193, .9)', line = list(color = 'rgba(152, 0, 0, .8)', width = 2)),
        text = ~paste("Country:", Country)) %>%
  layout(title = "Interactive Scatter Plot of GDP vs Child Deprivation")
```

## Conclusion:

Through these visualizations, our narrative bridges the gap between macroeconomic indicators and their human implications, offering a comprehensive view that connects economic health with child welfare outcomes. By understanding these relationships, policymakers, researchers, and the global community can better target interventions and policies to reduce child deprivation effectively.

This cohesive storyline, supported by diverse visual representations, not only informs but also empowers stakeholders to take informed action based on a clear understanding of where and how economic factors intersect with the lives of children around the world.
